---
title: "Multiple Imputation Analysis Using NHANES Data"
author: "JI YINGXIANG"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  beamer_presentation:
    theme: "Madrid"
    colortheme: "default"
    fonttheme: "professionalfonts"
    slide_level: 2
    keep_tex: true
fontsize: 10pt
header-includes:
  - \usepackage{booktabs}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{adjustbox}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 6.0,
  fig.height = 3.0,
  out.width = "0.88\\linewidth",
  dpi = 150
)
options(width = 60)

library(mice)
library(tidyverse)
library(knitr)
library(kableExtra)

kbl_fit <- function(x, caption = NULL, digits = 3) {
  x <- as.data.frame(x)
  for (j in seq_along(x)) if (is.numeric(x[[j]])) x[[j]] <- round(x[[j]], digits)
  kable(x, format = "latex", booktabs = TRUE, caption = caption, longtable = FALSE) |>
    kable_styling(latex_options = c("hold_position","scale_down"), font_size = 7)
}
```

## Objective
- Apply **Multiple Imputation (MI)** to a real dataset with missing data  
- Conduct regression analysis and pooled inference  
- Sensitivity analysis using different imputation numbers

## Data Source
- National Health and Nutrition Examination Survey (**NHANES**) 
- Teaching subset included in the `mice` package
- https://wwwn.cdc.gov/nchs/nhanes/

## Dataset Description
- Sample size: **25**
- Variables: age, bmi, hyp, chl

```{r data}
data(nhanes, package = "mice")
head(nhanes)
```

## Missing Data Pattern
```{r mdpattern, fig.height=2.6}
md.pattern(nhanes)
```

## Missing Data Assumption
- Missing values are assumed **Missing At Random (MAR)**  
- Missingness can be explained by observed covariates and study context

## Multiple Imputation Strategy
- Method: **MICE** with predictive mean matching  
- Main analysis uses **m = 20** imputations  
- A smaller number of imputations (**m = 5**) is used for robustness checking

## Convergence Diagnostics
```{r diag, fig.height=3.0}
set.seed(123)
imp20 <- mice(nhanes, m = 20, method = "pmm", printFlag = FALSE)
plot(imp20, c("bmi","chl"))
```

## Analysis Model
Linear regression model:
\[
\text{bmi} = \beta_0 + \beta_1 \text{age} + \beta_2 \text{hyp} + \varepsilon
\]

## Main Results (m = 20)
- This table reports the primary regression results after multiple imputation.
- Estimates are pooled across 20 imputed datasets using Rubin’s rules.

```{r fit20}
fit20 <- with(imp20, lm(bmi ~ age + hyp))
pooled20 <- pool(fit20)
tab20 <- summary(pooled20, conf.int = TRUE) |>
  as_tibble() |>
  select(term, estimate, std.error, statistic, df, p.value, `2.5 %`, `97.5 %`)
kbl_fit(tab20, caption = "Pooled regression results (m = 20)")
```

## Results with Fewer Imputations (m = 5)
- This analysis repeats the same model using fewer imputations.
- The goal is to assess whether conclusions depend strongly on the choice of *m*.

```{r fit5}
set.seed(123)
imp5 <- mice(nhanes, m = 5, method = "pmm", printFlag = FALSE)

fit5 <- with(imp5, lm(bmi ~ age + hyp))
pooled5 <- pool(fit5)
tab5 <- summary(pooled5, conf.int = TRUE) |>
  as_tibble() |>
  select(term, estimate, std.error, statistic, df, p.value, `2.5 %`, `97.5 %`)
kbl_fit(tab5, caption = "Pooled regression results (m = 5)")
```

## Comparison of Estimates
- Coefficient estimates and uncertainty are compared across *m = 20* and *m = 5*.
- Similar results suggest robustness to the number of imputations.

```{r compare}
cmp <- tab20 |>
  transmute(term,
            est_m20 = estimate, se_m20 = std.error, p_m20 = p.value,
            lo_m20 = `2.5 %`, hi_m20 = `97.5 %`) |>
  left_join(
    tab5 |>
      transmute(term,
                est_m5 = estimate, se_m5 = std.error, p_m5 = p.value,
                lo_m5 = `2.5 %`, hi_m5 = `97.5 %`),
    by = "term"
  )
kbl_fit(cmp, caption = "Comparison of pooled estimates by number of imputations")
```

## Interpretation
- Multiple imputation appropriately accounts for uncertainty due to missing data  
- Main findings are consistent across different numbers of imputations  
- Results do not appear sensitive to the choice of *m* in this example

## Gituhub
- https://github.com/toki0000/assignment3_mi_nhanes 


